---
title: "Data Wrangling Lecture 1"
author: "Felix Tran"
date: "September 18, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Importing FAS datasets

Filepaths for R. There are 2 ways to use them to import:

1. Absolute path -- Full address/pathfile for a file on a computer

2. Relative path -- Directions from your current destination to your file

Relative paths are better. Absolute paths only work on the specific computer
you use it on. Absolute paths also break when you move folders/files/etc. 
Relative paths work even when things change.

This is also a benefit of using Rprojects. Using an Rproject sets the working
directory to your Rproject. As long as you keep all your files in the Rproject,
you can open the Rproject anywhere and your relative paths for importing data
still work.

## Import FAS csv file
Import the FAS dataset
```{r}
litters_data <- read_csv('./data/FAS_litters.csv')
```

Look at column names
```{r}
names(litters_data)
```

R doesn't handle variable names with spaces/special symbols well. We will clean
them using the clean_names function from the janitor package.
```{r}
litters_data <- janitor::clean_names(litters_data)
names(litters_data)
```

Also read in the pups dataset and clean the variable names
```{r}
pups_data <- read_csv('./data/FAS_pups.csv')
names(pups_data)
pups_data <- janitor::clean_names(pups_data)
names(pups_data)
```

