---
title: "Data_manipulation"
author: "Felix Tran"
date: "September 20, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library('tidyverse')
litters_data <- read_csv('./data/FAS_litters.csv')
pups_data <- read_csv('./data/FAS_pups.csv')
litters_data <- janitor::clean_names(litters_data)
pups_data <- janitor::clean_names(pups_data)
```

## Select()

select() tells R to only keep the columns you want in the dataframe
```{r}
select(litters_data, litter_number, gd0_weight, pups_survive)
select(litters_data, gd_of_birth:pups_survive)
```

can also specify what you don't want to keep
```{r}
select(litters_data, -litter_number)
```

can also change col names in select() function. Alternatively can use rename()
function
```{r}
select(litters_data, group, litter_number, gest_day_0_weight = gd0_weight)
rename(litters_data, group_number = group)
```

select helpers i.e. everything() lets you choose var order by specifying 
variables in order, then call everything() to automatically select remaining
variables. For more of these search ?select_helpers
```{r}
select(litters_data, starts_with('gd'))
select(litters_data, litter_number, everything())
```

## Filtering observations
We can use logical operators like < > == != to filter observations
Using the filter() function makes things neat
Can use the filter() on multiple conditions with , (looks neater than & imo)
```{r}
filter(litters_data, gd0_weight < 25)
filter(litters_data, gd0_weight >= 25)
filter(litters_data, pups_born_alive == 8)
filter(litters_data, is.na(gd0_weight))
filter(litters_data, !is.na(gd0_weight))
filter(litters_data, group == "Low7" | group == "Low8")
filter(litters_data, group %in% c("Low7", "Low8"))
filter(litters_data, pups_born_alive %in% c(4,6,7), !is.na(gd0_weight))
```


## Mutate
Create variables. Can also do multiple operations in same function call.
Can do operations on a newly created variable in same function call.
```{r}
mutate(litters_data,
       wt_gain = gd18_weight - gd0_weight,
       group = tolower(group))
mutate(litters_data,
       wt_gain = gd18_weight - gd0_weight,
       wt_gain_squared = wt_gain * wt_gain)
```


## Arrange

arrange() sorts dataset by variable. Can sort on multiple variables
```{r}
arrange(litters_data, gd0_weight)
arrange(litters_data, pups_born_alive, gd0_weight)
arrange(litters_data, desc(pups_born_alive), gd0_weight)
```




