---
title: "Tidy_data"
author: "Felix Tran"
date: "September 25, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library('tidyverse')
```

## Tidy data
"long" data. One line = 1 observation.

Each column is 1 variable. It can be easy to combine information into 1 variable,
so we have to clean it.

## Import pulse dataset
```{r}
pulse_data <- haven::read_sas('./data/public_pulse_data.sas7bdat') %>% 
janitor::clean_names()
pulse_data
```

Not tidy yet

Use the gather() function
```{r}
pulse_data_tidy  <-
  gather(pulse_data, key = 'visit', value = bdi, bdi_score_bl:bdi_score_12m) 
pulse_data_tidy
```

Illustrate separate() function. Illustrates how we can separate the values of
visit into 3 variables (sep by the _), then remove 2 of the unneeded ones to
be left with just the info we need. 
```{r}
pulse_data_tidy <- 
  pulse_data_tidy %>%  
  separate(visit, into = c("bdi_str", "score_str", "visit"), sep = "_") %>% 
  select(-bdi_str, -score_str) %>% 
  mutate(visit = replace(visit, visit == "bl", "00m"))
pulse_data_tidy
```

## Exercise with litters data

```{r}
litters_data <-
  read_csv('./data/FAS_litters.csv', col_types = 'ccddiiii') %>% 
  janitor::clean_names()
litters_data

litters_data <-
  litters_data %>% 
  gather(key = 'gest_day', value = 'weight', gd0_weight:gd18_weight) %>% 
  select(litter_number, gest_day, weight) %>%  
  arrange(litter_number) %>% 
  mutate(gest_day = recode(gest_day, "gd0_weight" = 0, "gd18_weight" = 18))
litters_data

```


## Spread

Reverse function of gather(). Takes tidy data for R and transforms it into 
wide format data for human readibility.

The kable() function in the knitr package formats dataframe 2X2 into a neat
looking table for readibility.

```{r}
analysis_result = tibble(
  group = c("treatment", "treatment", "placebo", "placebo"),
  time = c("pre", "post", "pre", 'post'),
  mean = c(4, 8, 3.5, 4)
)

analysis_result

analysis_spread <-
  analysis_result %>% 
  spread(key = time, value = mean) %>% 
  knitr::kable()
analysis_spread
```

## Lord of the rings example
Read in the 3 tables 
```{r}
fellowship <- readxl::read_excel('./data/LotR_Words.xlsx', range = "B3:D6") %>% 
  mutate(movie = "fellowship_ring")
fellowship

towers <- readxl::read_excel('./data/LotR_Words.xlsx', range = "F3:H6") %>% 
  mutate(movie = "two_towers")
towers

king <- readxl::read_excel('./data/LotR_Words.xlsx', range = "J3:L6") %>% 
  mutate(movie = "return_king")
king

lotr <- bind_rows(fellowship, towers, king)
lotr
```

Not clean yet. Female and male should be recoded with a "gender" variable and then
a "word count" variable.
```{r}
lotr <-
  lotr %>% 
  gather(key = sex, value = word, Female:Male)

lotr
```


## Joining datasets
Using the left_join() function.

For clarity, joining the datasets is a separate step from reading in/modifying 
the individual datasets.
```{r}
pup_data = read_csv("./data/FAS_pups.csv", col_types = "ciiiii") %>%
  janitor::clean_names() %>%
  mutate(sex = recode(sex, `1` = "male", `2` = "female")) 

litter_data = read_csv("./data/FAS_litters.csv", col_types = "ccddiiii") %>%
  janitor::clean_names() %>%
  select(-pups_survive) %>%
  mutate(
    wt_gain = gd18_weight - gd0_weight,
    group = tolower(group))

fas_data = left_join(pup_data, litter_data, by = "litter_number")
fas_data
```

